Compilation With Program Synthesis

### 2023-08-22 new-paper/overview.tex
bibliography; a few sentences on layout polymorphism
code formatting

\begin{lstlisting}
void add1Proc(int x, loc r)
  { r :-> 0 }
  { y == x+1 ; r :-> y }
{ ?? }
\end{lstlisting}

@@

\begin{lstlisting}
add1Proc : Int -> Int;
add1Proc x := x + 1;
\end{lstlisting}

@@

There is a subtlety here: The layout of a subexpression inside of a
function definition might not be determined by the layouts that the
function definition is instantiated at. In this case, we must
explicitly specify a layout. This is done using the \verb|instantiate|
and \verb|lower| constructs. The former is for function calls, while
the latter is for all other types of values. These operations are
performed with the \verb|instantiate| and \verb|lower| language
constructs, respectively. For example, \verb|instantiate [A] B f x|
calls $f$ with the argument $x$, where the layout $A$ is used for the
argument and the layout $B$ is used for the result of the function
application. This is also written $\instantiateS{A,B}{f}(x)$. The
latter is primarily used in \autoref{sec:formal}.

@@

An example of an ADT definition and a corresponding layout definition
is given in \autoref{fig:List-def}. There is one unusual part of
the syntax in particular which requires further explanation: layout
type signatures. A layout definition consists a \textit{layout type
  signature} and a pattern match (much like a function definition),
with lists of SSL heaplets on the right-hand sides. A layout type
signature has a special form $A : \alpha \monic \layout[x]$. This says
that the layout $A$ is for the algebraic data type $\alpha$ and the
SSL output variable is named $x$. The full \Pika{} grammar is given in
the Appendices. %x~\ref{sec:grammar}.

\begin{figure}[t]
\begin{lstlisting}
data List := Nil | Cons Int List;

@@

\begin{lstlisting}
%generate singleton [Int] Sll

@@

\begin{lstlisting}
predicate singleton(int p, loc r) {
| true => { r :-> p ** (r+1) :-> 0 ** [r,2] }
}
\end{lstlisting}

@@

\begin{figure}[t]
\begin{lstlisting}
predicate sll(loc x) {
| x == 0 => { emp }
| not (x == 0) => { [x, 2] ** x :-> v ** (x+1) :-> nxt ** sll(nxt) }
}

@@

\begin{lstlisting}
mapAdd1_fn : List -> List;
mapAdd1_fn (Nil) := 0;
mapAdd1_fn (Cons head tail) := Cons (head + 1) (mapAdd1_fn tail);
\end{lstlisting}

@@

All that is left is to connect these two parts: the layouts and the
function definitions. We instruct a \SuSLik{} specification generator to
generate a \SuSLik{} specification from the \verb|mapAdd1_fn| function
using the \verb|Sll| layout:
%
\begin{lstlisting}
%generate mapAdd1_fn [Sll] Sll
\end{lstlisting}
%
The \verb|[Sll]| part of the directive tells the generator which
layouts are used for the arguments. In this case, the function only
has one argument and the \verb|Sll| layout is used. The \verb|Sll| at
the end specifies the layout for the result.

@@

\begin{lstlisting}
%generate mapAdd1 [Sll] Sll

@@

\begin{figure}
  \begin{lstlisting}
predicate filterLt9(loc x, loc r) {
| (x == 0) => { r == 0 ; emp }
| not (x == 0) && head < 9 =>
    { x :-> head ** (x+1) :-> tail ** [x,2] ** filterLt9(tail, r) }
| not (x == 0) && not (head < 9) =>
    { x :-> head ** (x+1) :-> tail ** [x,2] ** filterLt9(tail, y)
      ** r :-> head ** (r+1) :-> y ** [r,2] }
}

@@

\begin{lstlisting}
%generate filterLt9 [Sll] Sll

@@

\begin{lstlisting}
%generate maximum [Sll] Int

@@

\begin{lstlisting}
%generate leftList [TreeLayout] Sll

@@

We also demonstrate \verb|Ptr| types and the \verb|addr| builtin operation. This is somewhat similar to an \verb|@| pattern in Haskell,
but it is a low-level construct, indictating how pointers should be used, rather than a high-level construct.
Given a base type $\alpha$ (such as \verb|Int|), \verb|Ptr Int| is a type. This type is
passed around differently in the generated code. In particular, it is passed by reference.
\verb|addr| is the corresponding value-level introduction operation.
\begin{lstlisting}
%generate fold_List [Int, Sll] (Ptr Int)

@@

\begin{lstlisting}
predicate fold_List(int i, loc p_x ,loc __r) {
| (p_x == 0) => { __r :-> i }
| (not (p_x == 0)) => {
  p_x :-> x ** (p_x+1) :-> xs ** [p_x,2] **
  func f(p_x, __p_2, __r) **
  fold_List(i, xs, __p_2) **
  temploc __p_2 }
}
\end{lstlisting}
%
In the generated code, \verb|f| is applied to \verb|p_x| rather than
\verb|x| as a result of using \verb|Ptr| and \verb|addr|.

>>>>

\begin{lstlisting}[language=SynLang]
void add1Proc(int x, loc r)
  { r :-> 0 }
  { y == x+1 ; r :-> y }
{ ?? }
\end{lstlisting}

@@

\begin{lstlisting}[language=Pika]
add1Proc : Int -> Int;
add1Proc x := x + 1;
\end{lstlisting}

@@

There is a subtlety here: The layout of a subexpression inside of a
function definition might not be determined by the layouts that the
function definition is instantiated at. In this case, we must
explicitly specify a layout. This is done using the \verb|instantiate|
and \verb|lower| constructs. The former is for function calls, while
the latter is for all other types of values. These operations are
performed with the \verb|instantiate| and \verb|lower| language
constructs, respectively. For example, \verb|instantiate [A] B f x|
calls $f$ with the argument $x$, where the layout $A$ is used for the
argument and the layout $B$ is used for the result of the function
application. This is also written $\instantiateS{A,B}{f}(x)$. The
latter is primarily used in \autoref{sec:semantics}.

@@

An example of an ADT definition and a corresponding layout definition
is given in \autoref{fig:List-def}. There is one unusual part of
the syntax in particular which requires further explanation: layout
type signatures. A layout definition consists a \textit{layout type
  signature} and a pattern match (much like a function definition),
with lists of SSL heaplets on the right-hand sides. A layout type
signature has a special form $A : \alpha \monic \layout[x]$. This says
that the layout $A$ is for the algebraic data type $\alpha$ and the
SSL output variable is named $x$.
% The full \Pika{} grammar is given in
% the Appendices. %x~\ref{sec:grammar}.

\begin{figure}[t]
\begin{lstlisting}[language=Pika]
data List := Nil | Cons Int List;

@@

\begin{lstlisting}[language=Pika]
%generate singleton [Int] Sll

@@

\begin{lstlisting}[language=SynLang]
predicate singleton(int p, loc r) {
| true => { r :-> p ** (r+1) :-> 0 ** [r,2] }
}
\end{lstlisting}

@@

\begin{figure}[t]
  \begin{lstlisting}[language=SynLang]
predicate sll(loc x) {
| x == 0 => { emp }
| not (x == 0) => { [x, 2] ** x :-> v ** (x+1) :-> nxt ** sll(nxt) }
}

@@

\begin{lstlisting}[language=Pika]
mapAdd1_fn : List -> List;
mapAdd1_fn (Nil) := 0;
mapAdd1_fn (Cons head tail) := Cons (head + 1) (mapAdd1_fn tail);
\end{lstlisting}

@@

All that is left is to connect these two parts: the layouts and the
function definitions. We instruct a \SuSLik{} specification generator to
generate a \SuSLik{} specification from the \verb|mapAdd1_fn| function
using the \verb|Sll| layout:
%
\begin{lstlisting}[language=Pika]
%generate mapAdd1_fn [Sll] Sll
\end{lstlisting}
%
The \verb|[Sll]| part of the directive tells the generator which
layouts are used for the arguments. In this case, the function only
has one argument and the \verb|Sll| layout is used. The \verb|Sll| at
the end specifies the layout for the result.

@@

\begin{lstlisting}[language=Pika]
%generate mapAdd1 [Sll] Sll

@@

\begin{figure}
  \begin{lstlisting}[language=SynLang]
predicate filterLt9(loc x, loc r) {
| (x == 0) => { r == 0 ; emp }
| not (x == 0) && head < 9 =>
    { x :-> head ** (x+1) :-> tail ** [x,2] ** filterLt9(tail, r) }
| not (x == 0) && not (head < 9) =>
    { x :-> head ** (x+1) :-> tail ** [x,2] ** filterLt9(tail, y)
      ** r :-> head ** (r+1) :-> y ** [r,2] }
}

@@

\begin{lstlisting}[language=Pika]
%generate filterLt9 [Sll] Sll

@@

\begin{lstlisting}[language=Pika]
%generate maximum [Sll] Int

@@

\begin{lstlisting}[language=Pika]
%generate leftList [TreeLayout] Sll

@@

We also demonstrate \verb|Ptr| types and the \verb|addr| builtin operation. This is somewhat similar to an \verb|@| pattern in Haskell,
but it is a low-level construct, indictating how pointers should be used, rather than a high-level construct.
Given a base type $\alpha$ (such as \verb|Int|), \verb|Ptr Int| is a type. This type is
passed around differently in the generated code. In particular, it is passed by reference.
\verb|addr| is the corresponding value-level introduction operation.
\begin{lstlisting}[language=Pika]
%generate fold_List [Int, Sll] (Ptr Int)

@@

\begin{lstlisting}[language=SynLang]
predicate fold_List(int i, loc p_x ,loc __r) {
| (p_x == 0) => { __r :-> i }
| (not (p_x == 0)) => {
  p_x :-> x ** (p_x+1) :-> xs ** [p_x,2] **
  func f(p_x, __p_2, __r) **
  fold_List(i, xs, __p_2) **
  temploc __p_2 }
}
\end{lstlisting}
%
In the generated code, \verb|f| is applied to \verb|p_x| rather than
\verb|x| as a result of using \verb|Ptr| and \verb|addr|.

### 2023-08-22 new-paper/synth.tex
bibliography; a few sentences on layout polymorphism
Clarify section title
code formatting

\section{\synth{} Keyword: Synthesis Inside \Pika}
\label{sec:synth}

Next, we introduce a feature that allows us to synthesize a program from a special kind of \Pika{} type signature. This special kind of type signature allows us to directly access \SuSLik's pure propositions. These type signatures are called \synth-signatures and the syntax for invoking the \Pika{} synthesis on a \synth-signature is:
\[
  \synth\; f : \sigma
\]
\noindent
where $\sigma$ is a \synth-signature. The syntax of a \synth-signature is given in Figure~\ref{fig:synth-sig-syntax}.

@@

\begin{lstlisting}
Sll : layout[x](List)
Sll Nil := emp;
Sll (Cons head tail) :=
  x :-> head ** (x+1) :-> nxt
    ** Sll tail [nxt];
\end{lstlisting}

@@

\begin{lstlisting}
Sll : @(s : set) layout[x](List);
Sll Nil := s == {} ;; emp;
Sll (Cons head tail) :=
  s == {head} ++ s1
    ;;
  x :-> head ** (x+1) :-> nxt
    ** Sll tail @s1 [nxt];
\end{lstlisting}

@@

\begin{lstlisting}
synth append :
  s == s1 ++ s2
    ;;
  Sll @s1 -> Sll @s2 -> Sll @s;
\end{lstlisting}

>>>>

\section{\synth{} Keyword: Synthesis Inside \Pika{} 2}
\label{sec:synth}

Next, we introduce a feature that allows us to synthesise a program from a special kind of \Pika{} type signature. This special kind of type signature allows us to directly access \SuSLik's pure propositions. These type signatures are called \synth-signatures and the syntax for invoking the \Pika{} synthesis on a \synth-signature is:
\[
  \synth\; f : \sigma
\]
\noindent
where $\sigma$ is a \synth-signature. The syntax of a \synth-signature is given in Figure~\ref{fig:synth-sig-syntax}.

@@

\begin{lstlisting}[language=Pika]
Sll : List >-> layout[x]
Sll Nil := emp;
Sll (Cons head tail) :=
  x :-> head ** (x+1) :-> nxt
    ** Sll tail [nxt];
\end{lstlisting}

@@

\begin{lstlisting}[language=Pika]
Sll : @(s : set) List >-> layout[x];
Sll Nil := s == {} ;; emp;
Sll (Cons head tail) :=
  s == {head} ++ s1
    ;;
  x :-> head ** (x+1) :-> nxt
    ** Sll tail @s1 [nxt];
\end{lstlisting}

@@

\begin{lstlisting}[language=Pika]
synth append :
  s == s1 ++ s2
    ;;
  Sll @s1 -> Sll @s2 -> Sll @s;
\end{lstlisting}
